---
title: "Cleaning_data"
output:
  pdf_document: default
  html_document: default
date: "2023-06-08"
---

Libraries

```{r, message = FALSE, warning=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(childsds)
library(MatchIt)
library(vroom)
```

# Preparing data

## Sugary beverage intake

```{r}
soda_raw_data <- vroom(r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\data\raw\abcd_bkfs01.txt)", delim = "\t",col_select = c("subjectkey","interview_age", "sex", "eventname", "bkfs_soda")) 
#,"bkfs_sodafreq", "bkfs_sodaquan")) - these are extra vars we could use
#chose necessary columns only 

#bkfs_sodafreq - soft drinks freq/last week
#bkfs_sodaquan - soft drinks #/day

#rename cols
colnames(soda_raw_data)=c("ID", "age_months", "sex", "event", "soda_per_day")

#drop 1 row - it's notes, not data
soda_raw_data <- soda_raw_data[-1,]

#filter 2nd year data - cuz BKFS data is only available for 2nd year
summary(as.factor(soda_raw_data$event))
soda_raw_data = soda_raw_data[soda_raw_data$event == '2_year_follow_up_y_arm_1',]
#10414 subjects 

#drop weird 2 people who drink 777 sodas per day and NAs
soda_clean_data <- soda_raw_data[!soda_raw_data$soda_per_day == "777", ]
#10403
soda_clean_data <- soda_clean_data[complete.cases(soda_clean_data), ]
#10342

```


# Now pulling up BMI data and other vars from my old ABCD analysis

```{r}
#pulling data from my abcd analysis
pulled_data <- read.csv(r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\data\raw\SSRT_data.csv)")[ ,c('src_subject_id', 'anthro_bmi_calc', "anthro_height_calc", "anthro_weight_calc", "anthro_waist_cm", "race_ethnicity", "household.income.bl", "event_name")]
colnames(pulled_data)


#filter 1st year
#cause this is where they have full data on income, race-ethnicity
summary(as.factor(pulled_data$event_name))
pulled_data = pulled_data[pulled_data$event_name == 'baseline_year_1_arm_1',]
#11876 subs 


#renaming stupid cols
names(pulled_data)[names(pulled_data) == 'src_subject_id'] <- 'ID'
names(pulled_data)[names(pulled_data) == 'anthro_bmi_calc'] <- 'bmi'
names(pulled_data)[names(pulled_data) == 'anthro_height_calc'] <- 'height'
names(pulled_data)[names(pulled_data) == 'anthro_weight_calc'] <- 'weight'
names(pulled_data)[names(pulled_data) == 'household.income.bl'] <- 'household.income'
#11876 subs

#merge with soda data
#didn't include the event_name (8th column) 
soda_clean_data <- merge(soda_clean_data, pulled_data[,c(1:7)], by = "ID")
#drop NAs
summary(as.factor(soda_clean_data$bmi))
soda_clean_data <- soda_clean_data[complete.cases(soda_clean_data), ]
#9452 subs
```
# convert char columns into numeric
```{r}
lapply(soda_clean_data, class)
#all columns are chars

#convert char columns into numeric
colnames(soda_clean_data)

soda_clean_data[, c(2,5:9)] <- sapply(soda_clean_data[, c(2,5:9)], as.numeric)
```

#Add BMI%
```{r}

#names(dna)
soda_clean_data$age_years <- soda_clean_data$age_months/12

library(childsds)

soda_clean_data$bmi_perc <- childsds::sds(soda_clean_data$bmi,
                    age = soda_clean_data$age_years,
                    sex = soda_clean_data$sex, male = "M", female = "F",
                    ref = cdc.ref,
                    item = "bmi",
                    type = "perc")

soda_clean_data$bmi_perc <- soda_clean_data$bmi_perc*100
```


#Filter out underweights
#Add BMI% groups

```{r}
#filter out underweights

soda_clean_data <- soda_clean_data %>% filter(bmi_perc > 5)
summary(soda_clean_data$bmi_perc)
#8416 subs 

#add new column of bmi categories
soda_clean_data$bmi_perc_group <- cut(soda_clean_data$bmi_perc,
                          breaks=c(5, 85, 95, 100),
                          labels=c('Recommended', 'Overweight', 'Obese'))
summary(as.factor(soda_clean_data$bmi_perc_group))

#Recommended  Overweight       Obese 
#       6522        1064         830 
```



## Resting state correlation
```{r}
rest_corr_raw_data <- vroom(file = r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\data\raw\abcd_betnet02.txt)", delim = "\t")

#drop unneeded columns and row 1 (row 1 only has notes, no data)
colnames(rest_corr_raw_data)
rest_corr_raw_data <- rest_corr_raw_data[-1, -c(1:3, 5:8, 192, 193)]

#rename columns to better names
names(rest_corr_raw_data)[names(rest_corr_raw_data) == 'subjectkey'] <- 'ID'
names(rest_corr_raw_data)[names(rest_corr_raw_data) == 'eventname'] <- 'event'

#filter 2nd year
summary(as.factor(rest_corr_raw_data$event))
rest_corr_raw_data = rest_corr_raw_data[rest_corr_raw_data$event == '2_year_follow_up_y_arm_1',]

#drop unneccessary column (event) and nas
colnames(rest_corr_raw_data)
rest_corr_raw_data <- rest_corr_raw_data[, -2]
rest_corr_raw_data <- rest_corr_raw_data[complete.cases(rest_corr_raw_data), ]
#7626 subs


```

### Merging REST_CORR with the SODA umatched dataset
```{r}
soda_rest_unmatched <- merge(soda_clean_data, rest_corr_raw_data, by = "ID")
#7626 kids
```


## Subcortical

```{r}
data_raw_subcortical <- vroom(file = r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\data\raw\mrirscor02.txt)", delim = "\t")


#drop unneeded columns and row 1 (row 1 only has notes, no data)
colnames(data_raw_subcortical)
data_raw_subcortical <- data_raw_subcortical[-1, -c(1:3, 5:8, 270, 271)]
#262 vars

#rename columns to better names
names(data_raw_subcortical)[names(data_raw_subcortical) == 'subjectkey'] <- 'ID'
names(data_raw_subcortical)[names(data_raw_subcortical) == 'eventname'] <- 'event'

#filter 2nd year
summary(as.factor(data_raw_subcortical$event))
data_raw_subcortical = data_raw_subcortical[data_raw_subcortical$event == '2_year_follow_up_y_arm_1',]

#drop unneccessary column (event) and nas
colnames(data_raw_subcortical)

data_raw_subcortical <- data_raw_subcortical[, -2]
#261 cols
data_raw_subcortical <- data_raw_subcortical[complete.cases(data_raw_subcortical), ]
#7626 subs


```



```{r}
library(MatchIt)

#I will use the matchit function
#It requires 2 only - the group must be binary
#so I will create separate dfs with overweight and recommended folks

# I. REC BRAIN GROUP
# 1.1. rec group - dropping overweight
rec_brain_data <-subset(soda_rest_unmatched, soda_rest_unmatched$bmi_perc_group != "Overweight")

# 1.2 
rec_brain_data$bmi_perc_group <- droplevels(as.factor(rec_brain_data$bmi_perc_group))
summary(as.factor(rec_brain_data$bmi_perc_group))  
#Recommended       Obese 
#       4886         593

#1.3
rec_brain_data <- matchit(as.factor(bmi_perc_group) ~ age_years + sex + race_ethnicity, data = rec_brain_data)
# 1.4
rec_brain_data <- match.data(rec_brain_data)
summary(rec_brain_data$bmi_perc_group)
#Recommended       Obese 
#        593         593 

# II. BRAIN OV GROUP
# Now doing the same for the overweight group

#2.1. ov group - dropping overweight
ov_brain_data <- subset(soda_rest_unmatched, soda_rest_unmatched$bmi_perc_group != "Recommended")

#2.2
ov_brain_data$bmi_perc_group <- droplevels(as.factor(ov_brain_data$bmi_perc_group))
summary(as.factor(ov_brain_data$bmi_perc_group))

#2.3
ov_brain_data <- matchit(as.factor(bmi_perc_group) ~ age_years + sex, data = ov_brain_data)
summary(ov_brain_data)
#ov is some regression results
#so turning it into df
# 2.4 
ov_brain_data <- match.data(ov_brain_data)
summary(ov_brain_data$bmi_perc_group)
# Overweight      Obese 
#       593        593 

# III. BRAIN MERGING 

# 3.1. now drop obese from the ov, so that when I join 2 dfs, it won't have obese subs doubled
#now drop obese from the ov, so that when I join 2 dfs, it won't have obese subs doubled
ov_brain_data <- subset(ov_brain_data, ov_brain_data$bmi_perc_group != "Obese")
#still has obese group which makes our group non-binary
#3.2 
ov_brain_data$bmi_perc_group <- droplevels(as.factor(ov_brain_data$bmi_perc_group))
summary(as.factor(ov_brain_data$bmi_perc_group))  
#Overweight 
#       593

# 3.3 now merging dfs
soda_rest_corr <- bind_rows(rec_brain_data, ov_brain_data)
#199 cols
#1779 subs

# 3.4 have some weird columns created during the process, so deleteing them
soda_rest_corr <- soda_rest_corr[, c(-(197:199))]
#196
```


# Creating the same dataset for subcortical
```{r}
#I don't want to repeat all the bmi sorting staff
#So I'll just pull IDs from the resting state matched dataset
#And bind subcortical data instead

IDs <- data.frame(soda_rest_corr$ID)

#rename the column
colnames(IDs)
names(IDs)[names(IDs) == 'soda_rest_corr.ID'] <- 'ID'
soda_brain_matched <- merge(IDs, soda_clean_data, by = "ID")

#bind subcortical data
soda_subcortical <- merge(soda_brain_matched, data_raw_subcortical, by = "ID")

#checking
summary(as.factor(soda_subcortical$bmi_perc_group))

#Recommended  Overweight       Obese 
#        593         593         593 
```


# Plotting the distribution of sugar beverage intake

## Check if cols are numerical where needed
```{r}
lapply(soda_rest_corr, class)
lapply(soda_subcortical, class)
#all brain data columns are chars

#convert char columns into numeric
colnames(soda_rest_corr)
soda_rest_corr[, c(16:196)] <- sapply(soda_rest_corr[, c(16:196)], as.numeric)

colnames(soda_subcortical)
soda_subcortical[, c(16:274)] <- sapply(soda_subcortical[, c(16:274)], as.numeric)
```


## For beverage and resting state correlation
```{r}
hist(soda_rest_corr$soda_per_day)

#saving the plots into png files
summary(as.factor(soda_rest_corr$soda_per_day))

png(filename=r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\Scripts\0_Working\outputs\distribution_soda_per_day_rest.png)")
hist(soda_rest_corr$soda_per_day, main = "Distribution of soda per day \nn=1779 (BMI% groups matched)")
dev.off()

png(filename=r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\Scripts\0_Working\outputs\distribution_soda_per_day_subcort.png)")
hist(soda_subcortical$soda_per_day, main = "Distribution of sugary beverage intake frequency \nn=1779 (BMI% groups matched)")
dev.off()
```


```{r}
write.csv(soda_rest_corr, r"(C:\Users\Yanko\University of Wyoming\M2AENAD Lab - Documents\M2AENAD Lab - Documents\RESEARCH\ABCD\Yana_SSIB_2024\data\clean\rest_corr.csv)", row.names=FALSE)
write.csv(soda_subcortical, r"(C:\Users\Yanko\University of Wyoming\M2AENAD Lab - Documents\M2AENAD Lab - Documents\RESEARCH\ABCD\Yana_SSIB_2024\data\clean\subcortical.csv)", row.names=FALSE)
```

```{r}
summary(as.factor(soda_rest_corr$bmi_perc_group))
```



```{r}
#add mean centered age
#dna$mcage <- scale(dna$age, scale = FALSE)
```
