

# Libraries

```{r, message = FALSE, warning=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(childsds)
library(MatchIt)
library(vroom)
```


# Read in all the data at the same time
```{r}
path2data<-"/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/raw"
################################################################
soda_data <- vroom(paste(path2data,"abcd_bkfs01.txt", sep = "/"))
anthro_data <- vroom(paste(path2data,"abcd_ant01.txt",sep = "/"))
demo_data <- vroom(paste(path2data,"pdem02.txt",sep = "/"))
brain_data <- vroom(paste(path2data,"abcd_betnet02.txt",sep = "/"))
subcort_brain_data <-vroom(paste(path2data,"mrirscor02.txt",sep = "/"))
screen_data <- vroom(paste(path2data,"abcd_screen01.txt",sep = "/"))
med_data <- vroom(paste(path2data,"abcd_mx01.txt",sep = "/"))
family_data <- vroom(paste(path2data,"acspsw01.txt",sep = "/"))
imaging_site_data <- vroom(paste(path2data,"abcd_lt01.txt",sep = "/"))
#imaging_qc_data <- vroom(abcd_imgincl01.csv)

med_data <- med_data[-1, ]
```

```{r}
drops<-c("collection_id", "dataset_id",  "interview_date", "collection_title", "study_cohort_name")

```


```{r}
#summary(as.factor(anthro_data$eventname))
soda_df <- subset(soda_data, soda_data$eventname == "2_year_follow_up_y_arm_1")
soda_df<-soda_df[ , !(names(soda_df) %in% drops)]
anthro_df <- subset(anthro_data, anthro_data$eventname == "baseline_year_1_arm_1")
anthro_df<-anthro_df[ , !(names(anthro_df) %in% drops)]
demo_df <- subset(demo_data, demo_data$eventname == "baseline_year_1_arm_1")
demo_df<-demo_df[ , !(names(demo_df) %in% drops)]
brain_df <- subset(brain_data, brain_data$eventname == "baseline_year_1_arm_1")
brain_df<-brain_df[ , !(names(brain_df) %in% drops)]
subcort_brain_df <-subset(subcort_brain_data, subcort_brain_data$eventname == "baseline_year_1_arm_1")
subcort_brain_df<-subcort_brain_df[ , !(names(subcort_brain_df) %in% drops)]
screen_df <- subset(screen_data, screen_data$eventname == "baseline_year_1_arm_1")
screen_df<-screen_df[ , !(names(screen_df) %in% drops)]
med_df <- subset(med_data, med_data$eventname == "baseline_year_1_arm_1")
med_df<-med_df[ , !(names(med_df) %in% drops)]
imaging_site_df <- subset(imaging_site_data, imaging_site_data$eventname == "baseline_year_1_arm_1")
imaging_site_df<-imaging_site_df[ , !(names(imaging_site_df) %in% drops)]
```

```{r}
df<-join(anthro_df, demo_df)
df<-join(df, brain_df)
df<-join(df, subcort_brain_df)
df<-join(df, screen_df)
df<-join(df, med_df)
df<-join(df, imaging_site_df)

drops1<- c("eventname")
drops2 <- c("eventname","src_subject_id", "interview_age", "sex")
df<-df[ , !(names(df) %in% drops1)]
soda_df<-soda_df[ , !(names(soda_df) %in% drops2)]
df<-join(df,soda_df)
#11876
```
# First screen on obvious 
## Start n = 11875

```{r}
df <-subset(df, df$scrn_cpalsy != 0) #11188
df <- subset(df, df$scrn_hemorrhage != 0)#11184
df <- subset(df, df$scrn_medcond_other != 1)#10956
df <- subset(df, df$scrn_asd != 1)#10740
df <- subset(df, df$scrn_schiz != 1)#10738
df <- subset(df, df$scrn_tumor != 0)#10736
df <- subset(df, df$scrn_stroke != 0) #10735
df <- subset(df, df$scrn_aneurysm != 0)#10733
df <- subset(df, df$scrn_hemotoma != 0)#10733

```

Checking for ADHD, neurological condition, type 1 diabetes, muscle dystrophy, schizophrenia, autism
abcd_screen01
- scrn_cpalsy
-- n = 11867
- scrn_hemorrhage
-- n = 11863
- scrn_medcond_other
-- n = 11626
- scrn_asd
-- n = 11400
- scrn_schiz
-- 11397
- scrn_tumor
-- 11395
- scrn_stroke
- scrn_aneurysm
- scrn_hemorrhage
- scrn_hemotoma
- scrn_medcond_other
- scrn_schiz


abcd_mx01
- MusDys medhx_2l
- MS medhx_2m
- medhx_2k lead poisoning
- kidney disease medhx_2j
- cancer medhx_2e
- medhx_2g diabetes
- cere pals medhx_2f
- medhx_2o heart
```{r}
df<-subset(df, df$medhx_2l != 1)
df<-subset(df, df$medhx_2m != 1)
df<-subset(df, df$medhx_2k != 1)
df<-subset(df, df$medhx_2j != 1)
df<-subset(df, df$medhx_2e != 1)
df<-subset(df, df$medhx_2g != 1)
df<-subset(df, df$medhx_2f != 1)
df<-subset(df, df$medhx_2o != 1)#10297
```
output:
Joining by: collection_id, dataset_id, subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, collection_title
    0     1  NA's 
11486   384     5

```{r}
save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/screening65252.RData")
```

# demographics of interest
household income
1= Less than $5,000; 2=$5,000 through $11,999; 3=$12,000 through $15,999; 4=$16,000 through $24,999; 5=$25,000 through $34,999; 6=$35,000 through $49,999; 7=$50,000 through $74,999; 8= $75,000 through $99,999; 9=$100,000 through $199,999; 10=$200,000 and greater. 999 = Don't know  777 = Refuse to answer 
```{r}
df<-subset(df, df$demo_comb_income_v2 != "777")#9345
df<-subset(df, df$demo_comb_income_v2 != "999")#8947
```
#n demo data = 10321

```{r}
demos<-c("subjectkey", "sex", "interview_age", "demo_race_a_p___10", "demo_race_a_p___11", "demo_race_a_p___12",          "demo_race_a_p___13", "demo_race_a_p___14", "demo_race_a_p___15", "demo_race_a_p___16", "demo_race_a_p___17",         
 "demo_race_a_p___18", "demo_race_a_p___19", "demo_race_a_p___20", "demo_race_a_p___21", "demo_race_a_p___22",  "demo_race_a_p___23", "demo_race_a_p___24", "demo_comb_income_v2") 
demo_df<-demo_data[demos]
head(demo_df)
columns_to_sum <- c("demo_race_a_p___10", "demo_race_a_p___11", "demo_race_a_p___12",
                    "demo_race_a_p___13", "demo_race_a_p___14", "demo_race_a_p___15",
                    "demo_race_a_p___16", "demo_race_a_p___17", "demo_race_a_p___18",
                    "demo_race_a_p___19", "demo_race_a_p___20", "demo_race_a_p___21",
                    "demo_race_a_p___22", "demo_race_a_p___23", "demo_race_a_p___24")

# Convert specified columns to numeric
demo_df[columns_to_sum] <- lapply(demo_df[columns_to_sum], as.numeric)

# Calculate the sum across the specified columns for each row
demo_df$row_sum <- rowSums(demo_df[, columns_to_sum], na.rm = TRUE)
summary(as.factor(demo_df$row_sum))
demo_df$multiRacial <- ifelse(demo_df$row_sum > 1, 1, 0)
summary(as.factor(demo_df$multiRacial))
multi<-subset(demo_df, demo_df$multiRacial == 1)
multi$race<-multi$multiRacial*16
##################################################
single<-subset(demo_df, demo_df$multiRacial == 0)
#1 = white
single$race_white<-single$demo_race_a_p___10*1
#2 = black
single$race_black<-single$demo_race_a_p___11*2
#3 = native american
single$race_AmIn<-single$demo_race_a_p___12*3
#4 = native alaska
single$race_NaAl<-single$demo_race_a_p___13*4
#5 = native hawaiian
single$race_NaHi<-single$demo_race_a_p___14*5
#6 = guam
single$race_guam<-single$demo_race_a_p___15*6
#7 = samoa
single$race_samoa<-single$demo_race_a_p___16*7
#8 = PI
single$race_PI<-single$demo_race_a_p___17*8
#9 = IndianAsian
single$race_AIndian<-single$demo_race_a_p___18*9
#10 = Chinese
single$race_Chinese<-single$demo_race_a_p___19*10
#11 = Filipino
single$race_Filipino<-single$demo_race_a_p___20*11
#12 = Japanese
single$race_Japanese<-single$demo_race_a_p___21*12
#13 = Korean
single$race_Korean<-single$demo_race_a_p___22*13
#14 = Viet
single$race_Viet<-single$demo_race_a_p___23*14
#15 = OtherAsian
single$race_OAsian<-single$demo_race_a_p___24*15

sumagain<-c("race_white", "race_black", "race_AmIn", "race_NaAl", "race_NaHi", "race_guam", "race_samoa", "race_PI", "race_AIndian", "race_Filipino", "race_Chinese",  "race_Japanese", "race_Korean", "race_Viet","race_OAsian")
single$race <- rowSums(single[, sumagain], na.rm = TRUE)
```
  0    1    2    3    4    5 
 522 8691  968  123   16    1 
   0    1 
9213 1108 
Drop extra columns
```{r}
# Drop specific columns by name
single <- single %>%
  select(-demo_race_a_p___10, -demo_race_a_p___11, -demo_race_a_p___12, -demo_race_a_p___13, -demo_race_a_p___14, -demo_race_a_p___15, -demo_race_a_p___16, -demo_race_a_p___17, -demo_race_a_p___18, -demo_race_a_p___19, -demo_race_a_p___20, -demo_race_a_p___21, -demo_race_a_p___22, -demo_race_a_p___23, -demo_race_a_p___24, -row_sum, -multiRacial, -race_white, -race_black, -race_AmIn, -race_NaAl, -race_NaHi, -race_guam, -race_samoa, -race_PI, -race_AIndian, -race_Filipino, -race_Chinese, -race_Japanese, -race_Korean, -race_Viet, -race_OAsian)
```

```{r}
names(multi)
multi <- multi %>%
  select(-demo_race_a_p___10, -demo_race_a_p___11, -demo_race_a_p___12, -demo_race_a_p___13, -demo_race_a_p___14, -demo_race_a_p___15, -demo_race_a_p___16, -demo_race_a_p___17, -demo_race_a_p___18, -demo_race_a_p___19, -demo_race_a_p___20, -demo_race_a_p___21, -demo_race_a_p___22, -demo_race_a_p___23, -demo_race_a_p___24,  -row_sum, -multiRacial)
```

```{r}
DEMOS_df<-rbind(single, multi)
df<-join(df, DEMOS_df) #8947
summary(df$demo_comb_income_v2)
```
  1   10    2    3    4    5    6    7    8    9 
 367 1218  369  246  471  599  878 1426 1527 3220 
DEMOS_df 10321

```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/demos6525.RData")
```

```{r}
family_df<-family_data[,c("subjectkey","rel_family_id")]
```
```{r}
fams<-family_df$subjectkey

dfFAMS<-df[df$subjectkey %in% fams,]
dfFAMS<-join(dfFAMS, family_df)


dfnonFAMS<-df[!df$subjectkey %in% fams,]
start = 5000
nons = 5496
fakeFamily<-seq(start, start+nons , by=1)
dfnonFAMS$rel_family_id<-fakeFamily
dfnonFAMS$rel_family_id<-as.factor(dfnonFAMS$rel_family_id)

df<-rbind(dfnonFAMS, dfFAMS)
```



```{r}
df[,c("anthro_1_height_in", "anthro2heightin", "anthro3heightin", "anthroheightcalc", "anthroweight1lb", "anthroweight2lb", "anthroweight3lb", "anthroweightcalc", "anthro_waist_cm")][]<-lapply(df[,c("anthro_1_height_in", "anthro2heightin", "anthro3heightin", "anthroheightcalc", "anthroweight1lb", "anthroweight2lb", "anthroweight3lb", "anthroweightcalc", "anthro_waist_cm")], as.numeric)
```


```{r}
saveDF<-df

# Remove heights lower than 40 in and higher than 60 in
df<-subset(df, df$anthroheightcalc > 40) #8935 40 in is the below the 5% for girls age 9
```

```{r}
checkTalls<-subset(df$subjectkey, df$anthroheightcalc >= 60)
dfTallKeep<-df[df$subjectkey %in% tallTest$subjectkey,]

dfTalls<-anthro_data[anthro_data$subjectkey %in% checkTalls,]
summary(as.factor(anthro_data$eventname))
head(dfTalls)
inr<-c("subjectkey", "anthroheightcalc")
dfTalls_y0<-subset(dfTalls[inr], dfTalls$eventname == "baseline_year_1_arm_1")
dfTalls_y1<-subset(dfTalls[inr], dfTalls$eventname == "1_year_follow_up_y_arm_1")
colnames(dfTalls_y0) <- c("subjectkey", "anthroheightcalc0")
colnames(dfTalls_y1) <- c("subjectkey", "anthroheightcalc1")
names(dfTalls_y1)
tallTest<-join(dfTalls_y0, dfTalls_y1)
tallTest$growth<-as.numeric(tallTest$anthroheightcalc1) - as.numeric(tallTest$anthroheightcalc0)
summary(tallTest$growth)
tallTest<-subset(tallTest, tallTest$growth > 0)
tallTest<-subset(tallTest, tallTest$growth < 5 )
summary(as.numeric(tallTest$anthroheightcalc0))
```
```{r}
df<-subset(df, df$anthroheightcalc < 60) #8281 60 in is above 95% for boys age 9

df<-rbind(df, dfTallKeep) #8876
```


```{r}
#find weights that are significantly different (more 0.1 lbs)
t<-df
t$anthroweightAbsDiff12<-abs(t$anthroweight1lb - t$anthroweight2lb)
summary(t$anthroweightAbsDiff12)
t$diff12<-t$anthroweightAbsDiff12 > 0.1
summary(t$diff12)
t12OK<-subset(t, t$diff12 == FALSE)
t12OK$anthroweight1lb
t12OK$averageWTlb<-rowMeans(t12OK[,c("anthroweight1lb","anthroweight2lb")])
summary(t12OK$averageWTlb)

t12BAD<-subset(t, t$diff12 == TRUE)
t12BAD$diff13<-abs(t12BAD$anthroweight3lb - t12BAD$anthroweight1lb)
t12BAD$diff13bool<-t12BAD$diff13 > 0.1
t12BAD$diff23<-abs(t12BAD$anthroweight3lb - t12BAD$anthroweight2lb)
t12BAD$diff23bool<-t12BAD$diff23 > 0.1
summary(t12BAD$diff13bool)
summary(t12BAD$diff23bool)
head(t12BAD)

use23<-subset(t12BAD, t12BAD$diff23 == FALSE)
use23$averageWTlb<-rowMeans(use23[,c("anthroweight3lb","anthroweight2lb")])
use13<-subset(t12BAD, t12BAD$diff13 == FALSE)
use13$averageWTlb<-rowMeans(use13[,c("anthroweight3lb","anthroweight1lb")])
summary(use23$averageWTlb)
print(names(t12OK[,1000:1041]))
drops12<-c("anthroweightAbsDiff12", "diff12", "diff13",  "diff23", "diff13bool","diff23bool" )
drops23<-c("anthroweightAbsDiff12", "diff12",  "diff13", "diff23", "diff13bool", "diff23bool")
dropsOK12<-c("anthroweightAbsDiff12", "diff12")

t1<-t12OK[ , !(names(t12OK) %in% dropsOK12)]
t2<-use13[ , !(names(use13) %in% drops12)]
t3<-use23[ , !(names(use23) %in% drops23)]
## put back together

t<-rbind(t1,t2,t3)
df<-t #8625
summary(df$averageWTlb)
```


```{r}
dfRMLowWt <- subset(df, df$averageWTlb > 40) #8619
summary(dfRMLowWt$averageWTlb)
df<-dfRMLowWt
```

```{r}
df$BMI<-(df$averageWTlb/(df$anthroheightcalc)^2)*703
summary(df$BMI)
df$age_years<-as.numeric(df$interview_age)/12
summary(df$age_years)
```
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  2.148  15.891  17.558  18.658  20.400  52.825       5 

# Calculate BMI%
```{r}
df$bmi_perc <- childsds::sds(df$BMI,
                    age = df$age_years,
                    sex = df$sex, male = "M", female = "F",
                    ref = cdc.ref,
                    item = "bmi",
                    type = "perc")
summary(df$bmi_perc)
# filter out underweight children
df<-subset(df, df$bmi_perc>.05) #n = 8302
```


Check the motion
rsfmri_c_ngd_stnvols
Number of frames with FD < 0.2
Excluding subjects who have greater than 25% of total frames with FD >0.2
```{r}
df[,c("rsfmri_c_ngd_meanmotion","rsfmri_c_ngd_nvols", "rsfmri_c_ngd_stnvols")][]<-lapply(df[,c("rsfmri_c_ngd_meanmotion","rsfmri_c_ngd_nvols", "rsfmri_c_ngd_stnvols")], as.numeric)

df$FD25thresh<-df$rsfmri_c_ngd_nvols*.25

df$FDgreater02<-df$rsfmri_c_ngd_nvols - df$rsfmri_c_ngd_stnvols

df$bad0225<-df$FDgreater02 > df$FD25thresh
summary(df$bad0225) #3558 are bad 


failMotionCheck <- subset(df, df$bad0225 == TRUE)
df <- subset(df, df$bad0225 == FALSE) #4381
```

```{r}
df$bkfs_soda<-as.numeric(df$bkfs_soda)
summary(df$bkfs_soda)

df<-subset(df, df$bkfs_soda != "777") #n = 4002

df$ssb_group <- cut(df$bkfs_soda,
                          breaks=c(-Inf, 0, 2, Inf), 
                          labels=c('low', 'medium', 'high'))
summary(as.factor(df$ssb_group))

describeBy(df$bkfs_soda, df$ssb_group)
```


 low medium   high 
  1352   1809    841 
brain 160 - 589

```{r}
t<-df
names(df[,1000:1046])
interest<-c("subjectkey", "sex", "anthroheightcalc" , "averageWTlb", "demo_comb_income_v2", "BMI", "age_years", "bmi_perc", "ssb_group", "race", "rel_family_id", "bkfs_soda")
brain<-df[,160:589]
other<-df[,interest]
dftotal<-cbind(other, brain)
```



```{r}
#prepare datasets for easier data handling
baseline_anthro <- anthro_data[anthro_data$eventname == 'baseline_year_1_arm_1',]
y2_anthro <- anthro_data[anthro_data$eventname == '2_year_follow_up_y_arm_1',]

height_check_df <- screen_med_demo_df %>%
  inner_join(baseline_anthro, by = "subjectkey")
#height_check_df n=9872

# Create a column of height in cm by converting from inches to cm
height_check_df$height_cm <- as.numeric(height_check_df$anthroheightcalc) * 2.54


#subset those below 100 cm
short_subs <- height_check_df[height_check_df$height_cm < 100, ]
#short_subs n=12, of them 2 are NAs
summary(as.factor(short_subs$subjectkey))
colnames(short_subs)
#create a df to see the subjects' height at baseline and y2
short_subs_heights <- short_subs %>%
  select(subjectkey, height_baseline = anthroheightcalc) %>%
  left_join(
    y2_anthro %>% select(subjectkey, height_y2 = anthroheightcalc),
    by = "subjectkey"
  )

```
All heights are indeed abnormal, short subs must be excluded from the analysis

#Now weight QC

```{r}
#can't get this to run without creating all NAs<- GES
# Join weight data
weight_check_df <- screen_med_demo_df %>%
  inner_join(baseline_anthro, by = "subjectkey")
#weight_check_df n=9872

# Convert weight from lbs to kg
weight_check_df$weight_kg <- as.numeric(weight_check_df$anthroweightcalc) * 0.453592
weight_check_df$anthroweightcalc

# Subset those above 55 kg
heavy_subs <- weight_check_df[weight_check_df$weight_kg > 55, ]
summary(as.factor(heavy_subs$subjectkey))
#heavy_subs n = 687, 4 are NAs

# Create df to check weight at baseline and y2
heavy_subs_weights <- heavy_subs %>%
  select(subjectkey, weight_baseline = anthroweightcalc) %>%
  left_join(
    y2_anthro %>% select(subjectkey, weight_y2 = anthroweightcalc),
    by = "subjectkey"
  )

#687 subjects is too much to review
#start by dropping NAs
heavy_subs_weights <- na.omit(heavy_subs_weights)
#455

#now track the differences

# convert to kgs
heavy_subs_weights <- heavy_subs_weights %>%
  mutate(
    weight_baseline_kg = as.numeric(weight_baseline) * 0.453592,
    weight_y2_kg = as.numeric(weight_y2) * 0.453592
  )

gained_weight <- heavy_subs_weights %>%
  filter(weight_y2_kg > weight_baseline_kg)
#444
lost_weight <- heavy_subs_weights %>%
  filter(weight_y2_kg < weight_baseline_kg)
#10
#the differences are not too big
#conclusion - no exclusions needed due to abnormal weight 
```







```{r}
save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/combDF6525.RData")
```



#Create groups for SSB
If you give it an upper limit you don't need the part with the 777 it would remove those also if you have -Inf you might be picking up inaccurate negative soda intake, best to keep the bottom as 0. I just copied and pasted this above, but for the next iteration do this. 

If I keep the bottom at 0, it says the breaks are incorrect.
It's okay to use -Inf here since the minimum is 0.


```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/bmiDF6525.RData")
write.table(dftotal, "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/intDF.csv", sep = ",", row.names = F)
```

```{r}
df<-vroom("/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/intDF.csv")
```
```{r}
df<-df[complete.cases(df),] #3997
summary(as.factor(df$ssb_group))
```

# Create high vs low, and low vs. medium datasets
Since low will be the reference group, we want to match based on the reference group
# we don't need medium

```{r}
DFhigh_low<-subset(df, df$ssb_group != "medium") # n = 2192
summary(DFhigh_low$ssb_group)
DFhigh_low$ssb_group<-factor(DFhigh_low$ssb_group)
#this didn't drop the unused level
#DFhigh_low$ssb_group <- droplevels(DFhigh_low$ssb_group)
levels(DFhigh_low$ssb_group)
summary(DFhigh_low$ssb_group)
```
high  low 
 841 1350 

check siblings 
```{r}
sum(DFhigh_low$subjectkey %in% siblings$subjectkey)
```
#1832
If we only use subjects with relatives data, then we would have 1832 subjects only (before matching)
Thus, we will only exclude the proven siblings.

```{r}
sum(DFhigh_low$subjectkey %in% siblings$subjectkey) #1832

DFhigh_low_bkup <- DFhigh_low 
#DFhigh_low_bkup - pre-matched sample (b4 excluding for too much motion)
# DFhigh_low_bkup n=4580

not_in_siblings <- DFhigh_low %>%
  filter(!subjectkey %in% siblings$subjectkey)
#not_in_siblings - no family data available
#not_in_siblings n=2748

in_siblings <- siblings %>%
  filter(subjectkey %in% DFhigh_low$subjectkey)
#in_siblings n=1832 
#in_siblings - subjects with available family data

# Keep only 1 random subject per rel_family_id
set.seed(123)  # For reproducibility
no_siblings <- in_siblings %>%
  group_by(rel_family_id) %>%
  slice_sample(n = 1) %>%
  ungroup()
#no_siblings = sample with just 1 random kid per family
#no_siblings n=1682
data_for_no_siblings <- DFhigh_low_bkup %>%
  inner_join(no_siblings, by = "subjectkey")
DFhigh_low <- bind_rows(not_in_siblings, data_for_no_siblings)
#DFhigh_low_bkup- with cleaned siblings (the var DFhigh_low will be changed in further analysis)
#DFhigh_low n=4430
```
DFhigh_low n = 4430 (was 4580 before removing siblings)
```{r}
DFhigh_low$ssb_group <- factor(DFhigh_low$ssb_group, levels = c("high", "low"), labels = c(1, 0))

```


# Perform initial matching with null to get an idea of differences
```{r}
m.out0 <- matchit(ssb_group ~ age_years + bmi_perc, data = DFhigh_low,
                 method = NULL, distance = "glm")
```

```{r}
summary(m.out0)
```
# Perform matching using nearest neightbor (it works OK)
```{r}
m.out_highLow <- matchit(ssb_group ~ age_years + bmi_perc + demo_comb_income_v2, data = DFhigh_low,
                 method = "nearest", distance = "glm")

```


```{r}
summary(m.out_highLow)
plot(m.out_highLow)
```

# Get dataset from matched output
```{r}
highLow_DF <- match.data(m.out_highLow)
summary(m.out_highLow)
```
Sample Sizes:
          Control Treated
All          2645    1785
Matched      1785    1785
Unmatched     860       0
Discarded       0       0
total 3570
(this is  for subs with FD>0.9)
#FINAL 
Sample Sizes:
          Control Treated
All          2490    1665
Matched      1665    1665
Unmatched     825       0
Discarded       0       0

```{r}
highlow_DFbak <- highLow_DF
```
# Demographics
#1 = white
single$race_white<-single$demo_race_a_p___10*1
#2 = black
single$race_black<-single$demo_race_a_p___11*2
#3 = native american
single$race_AmIn<-single$demo_race_a_p___12*3
#4 = native alaska
single$race_NaAl<-single$demo_race_a_p___13*4
#5 = native hawaiian
single$race_NaHi<-single$demo_race_a_p___14*5
#6 = guam
single$race_guam<-single$demo_race_a_p___15*6
#7 = samoa
single$race_samoa<-single$demo_race_a_p___16*7
#8 = PI
single$race_PI<-single$demo_race_a_p___17*8
#9 = IndianAsian
single$race_AIndian<-single$demo_race_a_p___18*9
#10 = Chinese
single$race_Chinese<-single$demo_race_a_p___19*10
#11 = Filipino
single$race_Filipino<-single$demo_race_a_p___20*11
#12 = Japanese
single$race_Japanese<-single$demo_race_a_p___21*12
#13 = Korean
single$race_Korean<-single$demo_race_a_p___22*13
#14 = Viet
single$race_Viet<-single$demo_race_a_p___23*14
#15 = OtherAsian
single$race_OAsian<-single$demo_race_a_p___24*15

```{r}
sd(highlow_DFbak$age_years)
#2618/3700 #0.6227183
#2645/3570 #0.7408964
```
[1] 0.6227183
[1] 0.7408964
[1] 0.6203971
#Centering the data
```{r}
colnames(highLow_DF)

highLow_DF <- highLow_DF[,-c(433:437)]
highLow_DF[,c(5:8,12:443)]<-sapply(highLow_DF[,c(5:8,12:443)], as.numeric)
View(highLow_DF)

drops<-c("mrirscor02_id", "rsfmri_cor_ngd_scs_visitid","rsfmri_cor_ngd_scs_tr","rsfmri_cor_ngd_scs_numtrs",     
         "rsfmri_cor_ngd_scs_nvols", "rsfmri_cor_ngd_scs_stnvols", "rsfmri_cor_ngd_scs_stcgnvols", 
         "rsfmri_cor_ngd_scs_ntpoints", "rsfmri_cor_ngd_scs_meanmn", "rsfmri_cor_ngd_scs_maxmn", 
         "rsfmri_cor_ngd_scs_meantrans", "rsfmri_cor_ngd_scs_maxtrans", "rsfmri_cor_ngd_scs_meanrot", "rsfmri_cor_ngd_scs_maxrot",
         "distance", "weights", "subclass"    )
highLow_DF<-highLow_DF[ , !(names(highLow_DF) %in% drops)]

#Grace's original
highLow_DF[,c(5:8,12:428)]<-scale(highLow_DF[,c(5:8,12:428)], center = T, scale = F)
head(highLow_DF)

## Checking the center
round(colMeans(highLow_DF[,c(5:8,12:428)]),2)
sapply(highLow_DF[,c(5:8,12:428)], sd)

#mine
intDF[,c(1:8, 429)]<-sapply(intDF[,c(1:8, 429)], as.factor)
highLow_DF[,c(6, 9:428, 430:432)]<-sapply(highLow_DF[,c(6, 9:428, 430:432)], as.numeric)
highLow_DF[,c(6, 9:428, 430:432)]<-scale(highLow_DF[,c(6, 9:428, 430:432)], center = T, scale = F)
head(highLow_DF)
## Checking the center
round(colMeans(highLow_DF[,c(6, 9:428, 430:432)]),2) #all zeros
sapply(highLow_DF[,c(6, 9:428, 430:432)], sd)

```



```{r}
summary(highLow_DF$demo_comb_income_v2)
t.test(demo_comb_income_v2 ~ ssb_group, data = highLow_DF)

```

```{r}
library(ggplot2)
# Basic violin plot
ggplot(highlow_DFbak, aes(x = ssb_group, y = rsfmri_cor_ngd_sa_scs_thplh, color = as.factor(demo_comb_income_v2))) +
  geom_violin() +
  theme_minimal()
```

```{r}
library(ggpubr)
ch<-table(highlow_DFbak$ssb_group,highlow_DFbak$demo_comb_income_v2)
chisq.test(ch,correct=FALSE)
# Change color and fill
my_cols <- c("#0D0887FF", "#6A00A8FF", "#B12A90FF",
"#E16462FF", "#FCA636FF", "#F0F921FF")
ch<-as.data.frame(ch)
ch$Var2<-as.numeric(as.character(ch$Var2))
# Create the balloon plot
ggballoonplot(ch, x = "Var2", y = "Var1", size = "Freq", fill = "Freq") +
  scale_fill_gradientn(colors = my_cols) +
  labs(x = "Income Groups", y = "SSB Group", size = "Frequency", fill = "Frequency") +
  theme_minimal()

```
```{r}
# Scatterplot with a line of best fit
# Same, but with different colors and add regression lines
ggplot(highlow_DFbak, aes(x=as.numeric(as.character(demo_comb_income_v2)), y=rsfmri_c_ngd_cgc_ngd_cgc, color=ssb_group)) +
    geom_point(shape=1) +
    scale_colour_hue(l=50) + # Use a slightly darker palette than normal
    geom_smooth(method=lm) + theme_classic()

```

```{r}
install.packages("fastDummies")
library(fastDummies)
```


```{r}
highLow_DF_bkup3 <- highLow_DF
dummy_vars <- dummy_cols(highLow_DF, select_columns = "sex", remove_first_dummy = TRUE, remove_selected_columns = TRUE)
dummy_vars <- dummy_cols(dummy_vars, select_columns = "race", remove_first_dummy = FALSE, remove_selected_columns = TRUE)
levels(dummy_vars$ssb_group)
```


```{r}
write.table(dummy_vars, "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/matched6625_centered_dum.csv",  row.names = FALSE, sep = ",")

#write.table(dummy_vars, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data/matched111724_centered_dum.csv)",  row.names = FALSE, sep = ",")

```


# FOR BINARY REGRESSION, CREATE A DF WITH MOTION AND IMAGING SITE

# add in the motion for the brain data
```{r}
motion<-c("subjectkey","rsfmri_c_ngd_meanmotion")
brain_motion<-brain_data[,motion]
withMotion<-join(dummy_vars, brain_motion)
withMotion$rsfmri_c_ngd_meanmotionScaled<-scale(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), center = T, scale = F)
describe.by(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), withMotion$ssb_group)
```

add in the imaging sites
```{r}
withMotion_bkup <- withMotion
withMotion <- withMotion %>%
  left_join(imaging_site %>% select(subjectkey, site_id_l), by = "subjectkey")
```

#keep selected columns+ rename them
```{r}
colnames(withMotion)
binary_covars <- withMotion[,c()]#n=3330

binary_covars <- withMotion %>%
  select(subjectkey,
         meanmotionScaled = rsfmri_c_ngd_meanmotionScaled,
         imaging_site = site_id_l)
```


```{r}
write.table(withMotion, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data\matched021325_centered_dum.csv)",  row.names = FALSE, sep = ",")
write.table(binary_covars, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data\binary_covars.csv)",  row.names = FALSE, sep = ",")
```

```{r}
class(dummy_vars$demo_comb_income_v2)
```


```{r}
save.image(file = "data_cleaning_rerun.RData")
```






#I DID NOT RUN THE REST 

```{r}
#hist(highLow_DF$bkfs_soda, main = "Distribution of soda per day \nn=1568 (soda intake groups matched on age, bmi% and household income)")
```


```{r}
library(psych)
describeBy(df$rsfmri_cor_ngd_none_scs_cderh, df$ssb_group)
```

# Descriptives
```{r}
dim(highLow_DF)
```


```{r}
save.image(file = "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/finalDF6625.RData")
```

```{r}
load("~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/finalDF.RData")
```

```{r}
library(psych)
describeBy(highLow_DF$bkfs_soda, highLow_DF$ssb_group )
```

```{r}
library(ggplot2)
```

```{r}
names(highlow_DFbak)
```


```{r}
# Scatterplot with a line of best fit
# Same, but with different colors and add regression lines
ggplot(highlow_DFbak, aes(x=as.numeric(as.character(demo_comb_income_v2)), y=bmi_perc, color=ssb_group)) +
    geom_point(shape=1) +
    geom_smooth(method=lm) + theme_classic()+
    labs(
        x = "Combined Household Income",  # Custom x-axis label
        y = "BMI percentile"   # Custom y-axis label
    )+ scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

```

# adding in the motion for the brain data
```{r}
names(brain_data)
summary(as.numeric(brain_data$rsfmri_c_ngd_meanmotion))
motion<-c("subjectkey","rsfmri_c_ngd_meanmotion")
brain_motion<-brain_data[,motion]

withMotion<-join(dummy_vars, brain_motion)

withMotion$rsfmri_c_ngd_meanmotionScaled<-scale(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), center = T, scale = F)
describe.by(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), withMotion$ssb_group)
```


```{r}
library(psych)
describeBy(withMotion$rsfmri_c_ngd_meanmotion, withMotion$ssb_group)
```
```{r}
t.test(withMotion)
```


```{r}
#write.table(withMotion, "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/matched021325_centered_dum.csv",  row.names = FALSE, sep = ",")
write.table(withMotion, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data\matched021325_centered_dum.csv)",  row.names = FALSE, sep = ",")
```

#create a df of imaging sites to add to the binary regression
```{r}
final_subs <- withMotion$subjectkey
imaging_site_df <- imaging_site %>%
  filter(subjectkey %in% final_subs) %>%
  select(subjectkey, site_id_l) #n=3570
```

