

# Libraries

```{r, message = FALSE, warning=FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(childsds)
library(MatchIt)
library(vroom)
```


# Read in all the data at the same time
```{r}
path2data<-"/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/raw"
################################################################
soda_data <- vroom(paste(path2data,"abcd_bkfs01.txt", sep = "/"))
anthro_data <- vroom(paste(path2data,"abcd_ant01.txt",sep = "/"))
demo_data <- vroom(paste(path2data,"pdem02.txt",sep = "/"))
brain_data <- vroom(paste(path2data,"abcd_betnet02.txt",sep = "/"))
subcort_brain_data <-vroom(paste(path2data,"mrirscor02.txt",sep = "/"))
screen_data <- vroom(paste(path2data,"abcd_screen01.txt",sep = "/"))
med_data <- vroom(paste(path2data,"abcd_mx01.txt",sep = "/"))
family_data <- vroom(paste(path2data,"acspsw01.txt",sep = "/"))
imaging_site_data <- vroom(paste(path2data,"abcd_lt01.txt",sep = "/"))
#imaging_qc_data <- vroom(abcd_imgincl01.csv)

med_data <- med_data[-1, ]
```


ctrl + shift + c to comment out a block
<!-- # Create a quick data dict -->
<!-- ```{r} -->
<!-- DD<-cbind(head(soda_data,1), head(anthro_data, 1), head(demo_data,1), head(brain_data, 1), head(subcort_brain_data,1), head(anthro_data_2,1)) -->
<!-- DT <-data.frame(t(DD)) -->
<!-- DT$variable = row.names(DT)  -->
<!-- names(DT)[c(1, 2)] <- c("variable", "description") -->

<!-- #write.table(DT, "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/dataDict.csv", sep = ",", row.names = FALSE) -->

<!-- write.table(DT, r"(C:\Users\Yanko\University of Wyoming\M2AENAD Lab - Documents\M2AENAD Lab - Documents\RESEARCH\ABCD\Yana_SSIB_2024/data/dataDict.csv)", sep = ",", row.names = FALSE) -->

<!-- ``` -->


# First screen on obvious 
## Start n = 11875
Checking for ADHD, neurological condition, type 1 diabetes, muscle dystrophy, schizophrenia, autism
abcd_screen01
- scrn_cpalsy
-- n = 11867
- scrn_hemorrhage
-- n = 11863
- scrn_medcond_other
-- n = 11626
- scrn_asd
-- n = 11400
- scrn_schiz
-- 11397
- scrn_tumor
-- 11395
- scrn_stroke
- scrn_aneurysm
- scrn_hemorrhage
- scrn_hemotoma
- scrn_medcond_other
- scrn_schiz


```{r}
screen_data1 <-subset(screen_data, screen_data$scrn_cpalsy != 0)
screen_data2 <- subset(screen_data1, screen_data1$scrn_hemorrhage != 0)
screen_data3 <- subset(screen_data2, screen_data2$scrn_medcond_other != 1)
screen_data4 <- subset(screen_data3, screen_data3$scrn_asd != 1)
screen_data5 <- subset(screen_data4, screen_data4$scrn_schiz != 1)
screen_data6 <- subset(screen_data5, screen_data5$scrn_tumor != 0)
screen_data7 <- subset(screen_data6, screen_data6$scrn_stroke != 0)
screen_data8 <- subset(screen_data7, screen_data7$scrn_aneurysm != 0)
screen_data9 <- subset(screen_data8, screen_data8$scrn_hemotoma != 0)
screen_data10 <- subset(screen_data9, screen_data9$scrn_medcond_other != 1)
screen_df<-screen_data10 #11391
```
abcd_mx01
- MusDys medhx_2l
- MS medhx_2m
- medhx_2k lead poisoning
- kidney disease medhx_2j
- cancer medhx_2e
- medhx_2g diabetes
- cere pals medhx_2f
- medhx_2o heart
```{r}
med_data <- med_data[med_data$eventname == 'baseline_year_1_arm_1',]
screen_med_data<-join(med_data, screen_df)
####
summary(as.factor(screen_med_data$medhx_2o))
screen_med_data1<-subset(screen_med_data, screen_med_data$medhx_2l != 1)
screen_med_data2<-subset(screen_med_data1, screen_med_data1$medhx_2m != 1)
screen_med_data3<-subset(screen_med_data2, screen_med_data2$medhx_2k != 1)
screen_med_data4<-subset(screen_med_data3, screen_med_data3$medhx_2j != 1)
screen_med_data5<-subset(screen_med_data4, screen_med_data4$medhx_2e != 1)
screen_med_data6<-subset(screen_med_data5, screen_med_data5$medhx_2g != 1)
screen_med_data7<-subset(screen_med_data6, screen_med_data6$medhx_2f != 1)
screen_med_data8<-subset(screen_med_data7, screen_med_data7$medhx_2o != 1)
screen_med_df<-screen_med_data8 #11364
```
output:
Joining by: collection_id, dataset_id, subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, collection_title
    0     1  NA's 
11486   384     5

```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/screening6525.RData")
```

# demographics of interest
household income
1= Less than $5,000; 2=$5,000 through $11,999; 3=$12,000 through $15,999; 4=$16,000 through $24,999; 5=$25,000 through $34,999; 6=$35,000 through $49,999; 7=$50,000 through $74,999; 8= $75,000 through $99,999; 9=$100,000 through $199,999; 10=$200,000 and greater. 999 = Don't know  777 = Refuse to answer 
```{r}
demo_data<-subset(demo_data, demo_data$demo_comb_income_v2 != "777")
demo_data<-subset(demo_data, demo_data$demo_comb_income_v2 != "999")
demo_data <- demo_data[-1, ] #first columnrow is descriptions
```
#n demo data = 10321

```{r}
demos<-c("subjectkey", "sex", "interview_age", "demo_race_a_p___10", "demo_race_a_p___11", "demo_race_a_p___12",          "demo_race_a_p___13", "demo_race_a_p___14", "demo_race_a_p___15", "demo_race_a_p___16", "demo_race_a_p___17",         
 "demo_race_a_p___18", "demo_race_a_p___19", "demo_race_a_p___20", "demo_race_a_p___21", "demo_race_a_p___22",  "demo_race_a_p___23", "demo_race_a_p___24", "demo_comb_income_v2") 
demo_df<-demo_data[demos]
head(demo_df)
columns_to_sum <- c("demo_race_a_p___10", "demo_race_a_p___11", "demo_race_a_p___12",
                    "demo_race_a_p___13", "demo_race_a_p___14", "demo_race_a_p___15",
                    "demo_race_a_p___16", "demo_race_a_p___17", "demo_race_a_p___18",
                    "demo_race_a_p___19", "demo_race_a_p___20", "demo_race_a_p___21",
                    "demo_race_a_p___22", "demo_race_a_p___23", "demo_race_a_p___24")

# Convert specified columns to numeric
demo_df[columns_to_sum] <- lapply(demo_df[columns_to_sum], as.numeric)

# Calculate the sum across the specified columns for each row
demo_df$row_sum <- rowSums(demo_df[, columns_to_sum], na.rm = TRUE)
summary(as.factor(demo_df$row_sum))
demo_df$multiRacial <- ifelse(demo_df$row_sum > 1, 1, 0)
summary(as.factor(demo_df$multiRacial))
multi<-subset(demo_df, demo_df$multiRacial == 1)
multi$race<-multi$multiRacial*16
##################################################
single<-subset(demo_df, demo_df$multiRacial == 0)
#1 = white
single$race_white<-single$demo_race_a_p___10*1
#2 = black
single$race_black<-single$demo_race_a_p___11*2
#3 = native american
single$race_AmIn<-single$demo_race_a_p___12*3
#4 = native alaska
single$race_NaAl<-single$demo_race_a_p___13*4
#5 = native hawaiian
single$race_NaHi<-single$demo_race_a_p___14*5
#6 = guam
single$race_guam<-single$demo_race_a_p___15*6
#7 = samoa
single$race_samoa<-single$demo_race_a_p___16*7
#8 = PI
single$race_PI<-single$demo_race_a_p___17*8
#9 = IndianAsian
single$race_AIndian<-single$demo_race_a_p___18*9
#10 = Chinese
single$race_Chinese<-single$demo_race_a_p___19*10
#11 = Filipino
single$race_Filipino<-single$demo_race_a_p___20*11
#12 = Japanese
single$race_Japanese<-single$demo_race_a_p___21*12
#13 = Korean
single$race_Korean<-single$demo_race_a_p___22*13
#14 = Viet
single$race_Viet<-single$demo_race_a_p___23*14
#15 = OtherAsian
single$race_OAsian<-single$demo_race_a_p___24*15

sumagain<-c("race_white", "race_black", "race_AmIn", "race_NaAl", "race_NaHi", "race_guam", "race_samoa", "race_PI", "race_AIndian", "race_Filipino", "race_Chinese",  "race_Japanese", "race_Korean", "race_Viet","race_OAsian")
single$race <- rowSums(single[, sumagain], na.rm = TRUE)
```
  0    1    2    3    4    5 
 522 8691  968  123   16    1 
   0    1 
9213 1108 
Drop extra columns
```{r}
# Drop specific columns by name
single <- single %>%
  select(-demo_race_a_p___10, -demo_race_a_p___11, -demo_race_a_p___12, -demo_race_a_p___13, -demo_race_a_p___14, -demo_race_a_p___15, -demo_race_a_p___16, -demo_race_a_p___17, -demo_race_a_p___18, -demo_race_a_p___19, -demo_race_a_p___20, -demo_race_a_p___21, -demo_race_a_p___22, -demo_race_a_p___23, -demo_race_a_p___24, -row_sum, -multiRacial, -race_white, -race_black, -race_AmIn, -race_NaAl, -race_NaHi, -race_guam, -race_samoa, -race_PI, -race_AIndian, -race_Filipino, -race_Chinese, -race_Japanese, -race_Korean, -race_Viet, -race_OAsian)
```

```{r}
names(multi)
multi <- multi %>%
  select(-demo_race_a_p___10, -demo_race_a_p___11, -demo_race_a_p___12, -demo_race_a_p___13, -demo_race_a_p___14, -demo_race_a_p___15, -demo_race_a_p___16, -demo_race_a_p___17, -demo_race_a_p___18, -demo_race_a_p___19, -demo_race_a_p___20, -demo_race_a_p___21, -demo_race_a_p___22, -demo_race_a_p___23, -demo_race_a_p___24,  -row_sum, -multiRacial)
```

```{r}
DEMOS_df<-rbind(single, multi)
summary(as.factor(DEMOS_df$demo_comb_income_v2))
```
  1   10    2    3    4    5    6    7    8    9 
 367 1218  369  246  471  599  878 1426 1527 3220 
DEMOS_df 10321

```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/demos6525.RData")
```

```{r}
# Perform the join explicitly on "subjectkey"
screen_med_demo_df <- screen_med_df %>%
  inner_join(DEMOS_df, by = "subjectkey")
```
screen_med_demo_df 9873
#9873 was in Grace's version. too

# Screen for abnormal height and weight
```{r}
names(anthro_data)
anthroInterest<-c("subjectkey", "src_subject_id", "interview_age", "sex", "eventname", "anthro_1_height_in", "anthro2heightin", "anthro3heightin", "anthroheightcalc", "anthroweight1lb", "anthroweight2lb", "anthroweight3lb", "anthroweightcalc", "anthro_waist_cm")
anthro_data<-anthro_data[,anthroInterest]
head(anthro_data)
names(anthro_data)
anthro_data[,6:14][]<-lapply(anthro_data[,6:14], as.numeric)
head(anthro_data)
anthro_data$eventname<-as.factor(anthro_data$eventname)
anthro_data$sex<-as.factor(anthro_data$sex)
```


# First find the weights
```{r}
anthro_data$anthroweight1lb<-as.numeric(anthro_data$anthroweight1lb)
anthro_data$anthroweight2lb<-as.numeric(anthro_data$anthroweight2lb)
anthro_data$anthroweight3lb<-as.numeric(anthro_data$anthroweight3lb)
anthro_data$anthroheightcalc<-as.numeric(anthro_data$anthroheightcalc)
anthro_data$interview_age<-as.numeric(anthro_data$interview_age)
```

```{r}
summary(anthro_data$eventname)
anthroData0<-subset(anthro_data, anthro_data$eventname == "baseline_year_1_arm_1") #n = 11876
anthroData1<-subset(anthro_data, anthro_data$eventname == "1_year_follow_up_y_arm_1")
anthroData2<-subset(anthro_data, anthro_data$eventname == "2_year_follow_up_y_arm_1")

# Remove heights lower than 40 in and higher than 60 in
summary(anthroData01$anthro2heightin)
anthroData01<-subset(anthroData0, anthroData0$anthroheightcalc > 40) #11850 40 in is the below the 5% for girls age 9
anthroData01<-subset(anthroData0, anthroData0$anthroheightcalc < 60) #10955 60 in is above 95% for boys age 9

```


```{r}
#find weights that are significantly different (more 0.1 lbs)
anthroData01$anthroweightAbsDiff12<-abs(anthroData01$anthroweight1lb - anthroData01$anthroweight2lb)
summary(anthroData01$anthroweightAbsDiff12)
#> 0.1
anthroData01$anthroweightAbsDiff12 > 0.1
anthroData01$anthroweightGet3<- as.factor(anthroData01$anthroweightAbsDiff12 > 0.1)
summary(anthroData01$anthroweightGet3)

#remove implausabile weight lower than 40lb and greater than 120
WTOK<-subset(anthroData01, anthroData01$anthroweightGet3 == FALSE)
names(WTOK)
WTOK$averageWTlb<-rowMeans(WTOK[,10:11])
summary(WTOK$averageWTlb)
# weight less than 40lb
# weight greater than 120lb
WTOK<-subset(WTOK, WTOK$averageWTlb > 40)#9928
WTOK<-subset(WTOK, WTOK$averageWTlb < 120)#9435


checkWT<-subset(anthroData01, anthroData01$anthroweightGet3 == TRUE) #1015
checkWT$checkWT13<-abs(checkWT$anthroweight1lb - checkWT$anthroweight3lb)
checkWT$checkWT23<-abs(checkWT$anthroweight2lb - checkWT$anthroweight3lb)
summary(checkWT$checkWT13)
summary(checkWT$checkWT23)

checkWT$FLAG13<- checkWT$checkWT13 >0.1
summary(checkWT$FLAG13)

checkWT$FLAG23<- checkWT$checkWT23 >0.1
summary(checkWT$FLAG23)


checkWT<-subset(checkWT, checkWT$FLAG13 == FALSE | checkWT$FLAG23 == FALSE)#774
checkWTuse13<-subset(checkWT, checkWT$FLAG13 == FALSE) #291
checkWTuse13$averageWTlb<-rowMeans(checkWTuse13[c("anthroweight1lb", "anthroweight3lb" )])
names(checkWTuse13)
checkWTuse23<-subset(checkWT, checkWT$FLAG23 == FALSE) #488
checkWTuse23$averageWTlb<-rowMeans(checkWTuse23[c("anthroweight2lb", "anthroweight3lb" )])

checkWT2<-rbind(checkWTuse13, checkWTuse23)

dupes<-checkWT2[duplicated(checkWT2$subjectkey),]
dupes$subjectkey


for (sub in dupes$subjectkey){
  mu<-mean(checkWT2[checkWT2$subjectkey == sub,]$averageWTlb)
  checkWT2$averageWTlb[checkWT2$subjectkey == sub] <- mu
  print(checkWT2$averageWTlb[checkWT2$subjectkey == sub])


}

checkWT2<-checkWT2[!duplicated(checkWT2), ]#774
```

```{r}
drops <- c("checkWT13", "checkWT23", "FLAG13", "FLAG23")
checkWT3<-checkWT2[ , !(names(checkWT2) %in% drops)]

anthroData3<-rbind(checkWT3, WTOK) #10209
```


```{r}
summary(anthroData3$averageWTlb)
anthroData3<-subset(anthroData3, anthroData3$averageWTlb < 120) #10160
head(anthroData3)
```



```{r}
#prepare datasets for easier data handling
baseline_anthro <- anthro_data[anthro_data$eventname == 'baseline_year_1_arm_1',]
y2_anthro <- anthro_data[anthro_data$eventname == '2_year_follow_up_y_arm_1',]

height_check_df <- screen_med_demo_df %>%
  inner_join(baseline_anthro, by = "subjectkey")
#height_check_df n=9872

# Create a column of height in cm by converting from inches to cm
height_check_df$height_cm <- as.numeric(height_check_df$anthroheightcalc) * 2.54


#subset those below 100 cm
short_subs <- height_check_df[height_check_df$height_cm < 100, ]
#short_subs n=12, of them 2 are NAs
summary(as.factor(short_subs$subjectkey))
colnames(short_subs)
#create a df to see the subjects' height at baseline and y2
short_subs_heights <- short_subs %>%
  select(subjectkey, height_baseline = anthroheightcalc) %>%
  left_join(
    y2_anthro %>% select(subjectkey, height_y2 = anthroheightcalc),
    by = "subjectkey"
  )

```
All heights are indeed abnormal, short subs must be excluded from the analysis

#Now weight QC

```{r}
#can't get this to run without creating all NAs<- GES
# Join weight data
weight_check_df <- screen_med_demo_df %>%
  inner_join(baseline_anthro, by = "subjectkey")
#weight_check_df n=9872

# Convert weight from lbs to kg
weight_check_df$weight_kg <- as.numeric(weight_check_df$anthroweightcalc) * 0.453592
weight_check_df$anthroweightcalc

# Subset those above 55 kg
heavy_subs <- weight_check_df[weight_check_df$weight_kg > 55, ]
summary(as.factor(heavy_subs$subjectkey))
#heavy_subs n = 687, 4 are NAs

# Create df to check weight at baseline and y2
heavy_subs_weights <- heavy_subs %>%
  select(subjectkey, weight_baseline = anthroweightcalc) %>%
  left_join(
    y2_anthro %>% select(subjectkey, weight_y2 = anthroweightcalc),
    by = "subjectkey"
  )

#687 subjects is too much to review
#start by dropping NAs
heavy_subs_weights <- na.omit(heavy_subs_weights)
#455

#now track the differences

# convert to kgs
heavy_subs_weights <- heavy_subs_weights %>%
  mutate(
    weight_baseline_kg = as.numeric(weight_baseline) * 0.453592,
    weight_y2_kg = as.numeric(weight_y2) * 0.453592
  )

gained_weight <- heavy_subs_weights %>%
  filter(weight_y2_kg > weight_baseline_kg)
#444
lost_weight <- heavy_subs_weights %>%
  filter(weight_y2_kg < weight_baseline_kg)
#10
#the differences are not too big
#conclusion - no exclusions needed due to abnormal weight 
```

# Filter by event then drop event column
We are going to try to predict the SSB intake from the baseline data so we want baseline for the brain data
```{r}
soda_data <- soda_data[soda_data$eventname == '2_year_follow_up_y_arm_1',]
brain_data <- brain_data[brain_data$eventname == 'baseline_year_1_arm_1',]
demo_data <- demo_data[demo_data$eventname == 'baseline_year_1_arm_1',]
subcort_brain_data <- subcort_brain_data[subcort_brain_data$eventname == 'baseline_year_1_arm_1',]
anthro_data <- anthro_data[anthro_data$eventname == 'baseline_year_1_arm_1',]
baseline_imaging_site <- imaging_site_data[imaging_site_data$eventname == 'baseline_year_1_arm_1',]
baseline_siblings <- family_data[family_data$eventname == 'baseline_year_1_arm_1',]
```


```{r}
summary(as.factor(baseline_siblings$rel_group_id))
summary(as.factor(baseline_siblings$rel_family_id))
```


# Get columns of interest
```{r}
soda_interest<-c("subjectkey", "sex", "bkfs_soda")
anthro_interest<-c("subjectkey", "sex", "interview_age", "anthroheightcalc", "anthroweightcalc")
brain_interest<-c(4,8,23:191)
subcort_interest<-c(4,8,23:269)
siblings_interest <- c("subjectkey", "gender","rel_family_id")
imaging_site_interest<-c("subjectkey","site_id_l")
#subset
#colnames(brain_data)
soda_df<-soda_data[soda_interest]
anthro_df<-anthro_data[anthro_interest]
brain_df<-brain_data[,brain_interest]
subcort_df<-subcort_brain_data[,subcort_interest]
imaging_site <- baseline_imaging_site[imaging_site_interest]
siblings <- baseline_siblings[siblings_interest]
#n=4525 for siblings

```


# Join the datasets into a single dataframe
```{r}
df0<-join(soda_df, anthro_df)
df0.1<-join(df0,brain_df)
df0.2<-join(df0.1,subcort_df)

# Perform the join explicitly on "subjectkey"
df0.3 <- screen_med_demo_df %>%
  inner_join(df0.2, by = "subjectkey")

DF<-df0.3
```
#don't kow where this is from DEMOS n = 8981
#DEMOS_df n=10321

```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/combDF.RData")
```


#Create groups for SSB
If you give it an upper limit you don't need the part with the 777 it would remove those also if you have -Inf you might be picking up inaccurate negative soda intake, best to keep the bottom as 0. I just copied and pasted this above, but for the next iteration do this. 

If I keep the bottom at 0, it says the breaks are incorrect.
It's okay to use -Inf here since the minimum is 0.

```{r}
DF<-subset(DF, DF$bkfs_soda != "777") #n = 8926 (as in Grace's)
DF$bkfs_soda<-as.numeric(DF$bkfs_soda)
DF$ssb_group <- cut(DF$bkfs_soda,
                          breaks=c(-Inf, 0, 2, Inf), 
                          labels=c('low', 'medium', 'high'))
summary(as.factor(DF$ssb_group))
```
low medium   high 
  3021   3902   2003 
  
```{r}
describeBy(DF$bkfs_soda, DF$ssb_group)
```



# Calculate BMI and age in years
```{r}
DF$BMI<-(DF$anthroweightcalc/(DF$anthroheightcalc)^2)*703
DF$age_years<-DF$interview_age/12
summary(DF$BMI)
# Drop rows where column A has NA
DF <- DF[!is.na(DF$BMI), ] #n = 8921
```
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  2.148  15.891  17.558  18.658  20.400  52.825       5 

# Calculate BMI%
```{r}
DF$bmi_perc <- childsds::sds(DF$BMI,
                    age = DF$age_years,
                    sex = DF$sex, male = "M", female = "F",
                    ref = cdc.ref,
                    item = "bmi",
                    type = "perc")
summary(DF$bmi_perc)
# filter out underweight children
DF<-subset(DF, DF$bmi_perc>.05) #n = 8560 (as in Grace's)
```
  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 0.0000  0.3437  0.6513  0.6018  0.8943  0.9994 
 
```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/bmiDF.RData")
```

```{r}
names(DF)
#[1] "collection_id"                  "abcd_mx01_id"                   "dataset_id"                    
#  [4] "subjectkey"                     "src_subject_id"
# [166] "demo_comb_income_v2"            "race"                           "sex"                           
# [169] "bkfs_soda"                      "interview_age"                  "anthroheightcalc"  
#then comes the brain data and 
#[589] "ssb_group"                      "BMI"                            "age_years"                     
#[592] "bmi_perc"
intDF <- DF[, c(1:5, 166:592)]
#n = 8560 (like in Grace's)
```

#Drop missing values
```{r}
intDF<-intDF[complete.cases(intDF), ] #n = 8116 (like in Grace's)
#drop short_subs
intDF_filtered <- intDF %>%
  filter(!subjectkey %in% short_subs$subjectkey)
```

#Find the class of all columns and then cast columns as factors or numeric as needed
Why the factoring does not work? It keeps appearing as a character.

```{r}
colnames(intDF)
intDF[,c(1:8, 428)]<-sapply(intDF[,c(1:8, 428)], as.factor)
intDF[,c(9:427, 429:432)]<-sapply(intDF[,c(9:427, 429:432)], as.numeric)
summary(as.factor(intDF$ssb_group))
```
  1    2    3 
2730 3536 1850 
low medium high
total intDF n =8 116 
```{r}
#save.image(file = "/Users/gracer/Library/CloudStorage/OneDrive-UniversityofWyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/bmiDF.RData")
```

# Create high vs low, and low vs. medium datasets
Since low will be the reference group, we want to match based on the reference group
# we don't need medium

```{r}
DFhigh_low<-subset(intDF, intDF$ssb_group != "2") # n = 4580
DFhigh_low$ssb_group<-factor(DFhigh_low$ssb_group)
#this didn't drop the unused level
#DFhigh_low$ssb_group <- droplevels(DFhigh_low$ssb_group)
levels(DFhigh_low$ssb_group)
summary(DFhigh_low$ssb_group)
```
1    3 
2730 1850 

check siblings 
```{r}
sum(DFhigh_low$subjectkey %in% siblings$subjectkey)
```
#1832
If we only use subjects with relatives data, then we would have 1832 subjects only (before matching)
Thus, we will only exclude the proven siblings.

```{r}
sum(DFhigh_low$subjectkey %in% siblings$subjectkey) #1832

DFhigh_low_bkup <- DFhigh_low 
#DFhigh_low_bkup - pre-matched sample (b4 excluding for too much motion)
# DFhigh_low_bkup n=4580

not_in_siblings <- DFhigh_low %>%
  filter(!subjectkey %in% siblings$subjectkey)
#not_in_siblings - no family data available
#not_in_siblings n=2748

in_siblings <- siblings %>%
  filter(subjectkey %in% DFhigh_low$subjectkey)
#in_siblings n=1832 
#in_siblings - subjects with available family data

# Keep only 1 random subject per rel_family_id
set.seed(123)  # For reproducibility
no_siblings <- in_siblings %>%
  group_by(rel_family_id) %>%
  slice_sample(n = 1) %>%
  ungroup()
#no_siblings = sample with just 1 random kid per family
#no_siblings n=1682
data_for_no_siblings <- DFhigh_low_bkup %>%
  inner_join(no_siblings, by = "subjectkey")
DFhigh_low <- bind_rows(not_in_siblings, data_for_no_siblings)
#DFhigh_low_bkup- with cleaned siblings (the var DFhigh_low will be changed in further analysis)
#DFhigh_low n=4430
```
DFhigh_low n = 4430 (was 4580 before removing siblings)




#2067

# Perform initial matching with null to get an idea of differences
```{r}
m.out0 <- matchit(ssb_group ~ age_years + bmi_perc, data = DFhigh_low,
                 method = NULL, distance = "glm")
#m.out0.1 <- matchit(ssb_group ~ age_years + bmi_perc, data = DFmedium_low,
#                 method = NULL, distance = "glm")

```

```{r}
summary(m.out0)
```
# Perform matching using nearest neightbor (it works OK)
```{r}
m.out_highLow <- matchit(ssb_group ~ age_years + bmi_perc + demo_comb_income_v2, data = DFhigh_low,
                 method = "nearest", distance = "glm")

#m.out_mediumLow <- matchit(ssb_group ~ age_years + bmi_perc + household.income.bl, data = DFmedium_low,
#                 method = "nearest", distance = "glm")

```


```{r}
summary(m.out_highLow)
plot(m.out_highLow)
```

# Get dataset from matched output
```{r}
highLow_DF <- match.data(m.out_highLow)
summary(m.out_highLow)
```
Sample Sizes:
          Control Treated
All          2645    1785
Matched      1785    1785
Unmatched     860       0
Discarded       0       0
total 3570
(this is  for subs with FD>0.9)
#FINAL 
Sample Sizes:
          Control Treated
All          2490    1665
Matched      1665    1665
Unmatched     825       0
Discarded       0       0

```{r}
highlow_DFbak <- highLow_DF
names(highlow_DFbak)
highlow_DFbak
```
# Demographics
#1 = white
single$race_white<-single$demo_race_a_p___10*1
#2 = black
single$race_black<-single$demo_race_a_p___11*2
#3 = native american
single$race_AmIn<-single$demo_race_a_p___12*3
#4 = native alaska
single$race_NaAl<-single$demo_race_a_p___13*4
#5 = native hawaiian
single$race_NaHi<-single$demo_race_a_p___14*5
#6 = guam
single$race_guam<-single$demo_race_a_p___15*6
#7 = samoa
single$race_samoa<-single$demo_race_a_p___16*7
#8 = PI
single$race_PI<-single$demo_race_a_p___17*8
#9 = IndianAsian
single$race_AIndian<-single$demo_race_a_p___18*9
#10 = Chinese
single$race_Chinese<-single$demo_race_a_p___19*10
#11 = Filipino
single$race_Filipino<-single$demo_race_a_p___20*11
#12 = Japanese
single$race_Japanese<-single$demo_race_a_p___21*12
#13 = Korean
single$race_Korean<-single$demo_race_a_p___22*13
#14 = Viet
single$race_Viet<-single$demo_race_a_p___23*14
#15 = OtherAsian
single$race_OAsian<-single$demo_race_a_p___24*15

```{r}
sd(highlow_DFbak$age_years)
#2618/3700 #0.6227183
#2645/3570 #0.7408964
```
[1] 0.6227183
[1] 0.7408964
[1] 0.6203971
#Centering the data
```{r}
colnames(intDF)
#  [1] "collection_id"                  "abcd_mx01_id"                   "dataset_id"                    
#  [4] "subjectkey"                     "src_subject_id"                 "demo_comb_income_v2"           
#  [7] "race"                           "sex"                            "bkfs_soda"                     
# [10] "interview_age"                  "anthroheightcalc"               "anthroweightcalc"              
# [13] "rsfmri_c_ngd_ad_ngd_ad"         "rsfmri_c_ngd_ad_ngd_cgc"        "rsfmri_c_ngd_ad_ngd_ca"   
# [427] "rsfmri_cor_ngd_vs_scs_aarh"     "rsfmri_cor_ngd_vs_scs_vtdcrh"   "ssb_group"                     
# [430] "BMI"                            "age_years"                      "bmi_perc" 
colnames(highLow_DF)
highLow_DF <- highLow_DF[,-c(433:437)]
#Grace's original
#intDF[,c(1:8, 428)]<-sapply(intDF[,c(1:8, 428)], as.factor)
#highLow_DF[,c(6, 9:427, 429:432)]<-sapply(highLow_DF[,c(6, 9:427, 429:432)], as.numeric)
#highLow_DF[,c(6, 9:427, 429:432)]<-scale(highLow_DF[,c(6, 9:427, 429:432)], center = T, scale = F)
#head(highLow_DF)
## Checking the center
#round(colMeans(highLow_DF[,c(6, 9:427, 429:432)]),2)
#sapply(highLow_DF[,c(6, 9:427, 429:432)], sd)

#mine
intDF[,c(1:8, 429)]<-sapply(intDF[,c(1:8, 429)], as.factor)
highLow_DF[,c(6, 9:428, 430:432)]<-sapply(highLow_DF[,c(6, 9:428, 430:432)], as.numeric)
highLow_DF[,c(6, 9:428, 430:432)]<-scale(highLow_DF[,c(6, 9:428, 430:432)], center = T, scale = F)
head(highLow_DF)
## Checking the center
round(colMeans(highLow_DF[,c(6, 9:428, 430:432)]),2) #all zeros
sapply(highLow_DF[,c(6, 9:428, 430:432)], sd)

```

```{r}
describeBy(highLow_DF$interview_age, highLow_DF$ssb_group)
```


```{r}
summary(highLow_DF$demo_comb_income_v2)
t.test(demo_comb_income_v2 ~ ssb_group, data = highLow_DF)

library(psych)
#describeBy(highlow_DFbak$ssb_group)
```
```{r}
library(ggplot2)
# Basic violin plot
ggplot(highlow_DFbak, aes(x = ssb_group, y = rsfmri_cor_ngd_sa_scs_thplh, color = demo_comb_income_v2)) +
  geom_violin() +
  theme_minimal()
```

```{r}
library(ggpubr)
ch<-table(highlow_DFbak$ssb_group,highlow_DFbak$demo_comb_income_v2)
chisq.test(ch,correct=FALSE)
# Change color and fill
my_cols <- c("#0D0887FF", "#6A00A8FF", "#B12A90FF",
"#E16462FF", "#FCA636FF", "#F0F921FF")
ch<-as.data.frame(ch)
ch$Var2<-as.numeric(as.character(ch$Var2))
# Create the balloon plot
ggballoonplot(ch, x = "Var2", y = "Var1", size = "Freq", fill = "Freq") +
  scale_fill_gradientn(colors = my_cols) +
  labs(x = "Income Groups", y = "SSB Group", size = "Frequency", fill = "Frequency") +
  theme_minimal()

```
```{r}
# Scatterplot with a line of best fit
# Same, but with different colors and add regression lines
ggplot(highlow_DFbak, aes(x=as.numeric(as.character(demo_comb_income_v2)), y=rsfmri_c_ngd_cgc_ngd_cgc, color=ssb_group)) +
    geom_point(shape=1) +
    scale_colour_hue(l=50) + # Use a slightly darker palette than normal
    geom_smooth(method=lm) + theme_classic()

```

```{r}
install.packages("fastDummies")
library(fastDummies)
```


```{r}
highLow_DF_bkup3 <- highLow_DF
dummy_vars <- dummy_cols(highLow_DF, select_columns = "sex", remove_first_dummy = TRUE, remove_selected_columns = TRUE)
View(dummy_vars)
dummy_vars <- dummy_cols(dummy_vars, select_columns = "race", remove_first_dummy = FALSE, remove_selected_columns = TRUE)
levels(dummy_vars$ssb_group)
```


```{r}
#write.table(dummy_vars, "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/matched111724_centered_dum.csv",  row.names = FALSE, sep = ",")

write.table(dummy_vars, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data/matched111724_centered_dum.csv)",  row.names = FALSE, sep = ",")

```


# FOR BINARY REGRESSION, CREATE A DF WITH MOTION AND IMAGING SITE

# add in the motion for the brain data
```{r}
motion<-c("subjectkey","rsfmri_c_ngd_meanmotion")
brain_motion<-brain_data[,motion]
withMotion<-join(dummy_vars, brain_motion)
withMotion$rsfmri_c_ngd_meanmotionScaled<-scale(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), center = T, scale = F)
describe.by(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), withMotion$ssb_group)
```

add in the imaging sites
```{r}
withMotion_bkup <- withMotion
withMotion <- withMotion %>%
  left_join(imaging_site %>% select(subjectkey, site_id_l), by = "subjectkey")
```

#keep selected columns+ rename them
```{r}
colnames(withMotion)
binary_covars <- withMotion[,c()]#n=3330

binary_covars <- withMotion %>%
  select(subjectkey,
         meanmotionScaled = rsfmri_c_ngd_meanmotionScaled,
         imaging_site = site_id_l)
```


```{r}
write.table(withMotion, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data\matched021325_centered_dum.csv)",  row.names = FALSE, sep = ",")
write.table(binary_covars, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data\binary_covars.csv)",  row.names = FALSE, sep = ",")
```

```{r}
class(dummy_vars$demo_comb_income_v2)
```


```{r}
save.image(file = "data_cleaning_rerun.RData")
```






#I DID NOT RUN THE REST 

```{r}
#hist(highLow_DF$bkfs_soda, main = "Distribution of soda per day \nn=1568 (soda intake groups matched on age, bmi% and household income)")
```


```{r}
library(psych)
describeBy(df$rsfmri_cor_ngd_none_scs_cderh, df$ssb_group)
```

# Descriptives
```{r}
dim(highLow_DF)
```


```{r}
#save.image(file = "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/finalDF.RData")
```

```{r}
load("~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/finalDF.RData")
```

```{r}
library(psych)
describeBy(highLow_DF$bkfs_soda, highLow_DF$ssb_group )
```

```{r}
library(ggplot2)
```

```{r}
names(highlow_DFbak)
```


```{r}
# Scatterplot with a line of best fit
# Same, but with different colors and add regression lines
ggplot(highlow_DFbak, aes(x=as.numeric(as.character(demo_comb_income_v2)), y=bmi_perc, color=ssb_group)) +
    geom_point(shape=1) +
    geom_smooth(method=lm) + theme_classic()+
    labs(
        x = "Combined Household Income",  # Custom x-axis label
        y = "BMI percentile"   # Custom y-axis label
    )+ scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

```

# adding in the motion for the brain data
```{r}
names(brain_data)
summary(as.numeric(brain_data$rsfmri_c_ngd_meanmotion))
motion<-c("subjectkey","rsfmri_c_ngd_meanmotion")
brain_motion<-brain_data[,motion]

withMotion<-join(dummy_vars, brain_motion)

withMotion$rsfmri_c_ngd_meanmotionScaled<-scale(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), center = T, scale = F)
describe.by(as.numeric(withMotion$rsfmri_c_ngd_meanmotion), withMotion$ssb_group)
```


```{r}
library(psych)
describeBy(withMotion$rsfmri_c_ngd_meanmotion, withMotion$ssb_group)
```
```{r}
t.test(withMotion)
```


```{r}
#write.table(withMotion, "~/OneDrive - University of Wyoming/0. Lab/M2AENAD Lab - Documents/RESEARCH/ABCD/Yana_SSIB_2024/data/matched021325_centered_dum.csv",  row.names = FALSE, sep = ",")
write.table(withMotion, r"(C:\Users\Yanko\OneDrive - University of Wyoming\Desktop - Copy\Lab\SSIB 2024\0 PAPER\peer review\data\matched021325_centered_dum.csv)",  row.names = FALSE, sep = ",")
```

#create a df of imaging sites to add to the binary regression
```{r}
final_subs <- withMotion$subjectkey
imaging_site_df <- imaging_site %>%
  filter(subjectkey %in% final_subs) %>%
  select(subjectkey, site_id_l) #n=3570
```

